R Notebook

### 2020-02-20 MS.Rmd
Bioclim

```{r echo=FALSE, include=FALSE}
#adding 2008 and 2018 CPP values 
cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
gdd400_all_new<-left_join(gdd400_all, cpp_year, by="pop")
gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)

@@

We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions. Insect phenology studies often estimate generation times using growing degree day (GDD) models. Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 GDD. We used climate records over the past 20 years from our sampling locations to calculate a yearly GDD-based "deadline" for diapause at each location: the date after which eggs laid would not have enough growing degree days left in the year to complete another generation.

@@

```{r echo=FALSE}
kable(rbind(prettify(anova(model1)), prettify(anova(model2)), prettify(anova(model3)), prettify(anova(model4))), digits=3)  
```
```{r echo=FALSE}
prev_year_plots
```

```{r echo=FALSE}
kable(prettify(anova(all_prevyear_model))) 
kable(prettify(anova(all_previnterval_model)))

@@

The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process. Based on the distribution of percent differences from the initial counts (Figure S1A), we considered oviposition papers with more than 100% or less than 60% of the initial count to be less reliable data points; these papers made up less than 10% of both individual papers and total eggs quantified in the experiment. Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves. These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations (Figure S2). The final photoperiod response curves are shown in Figure S3.

@@

```{r, echo=FALSE}
DI_plot 
```
```{r echo=FALSE}
gdd_deadline_plots
```
```{r echo=FALSE}
plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE), labels="AUTO")

```
```{r echo=FALSE, warning=FALSE}
gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd400_mean, fill=country, shape=interval)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Period") + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) +
  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Period") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=12)) + xlab("Latitude") + ylab("Mean GDD Deadline")

>>>>

```{r echo=FALSE, include=FALSE}
#adding 2008 and 2018 CPP values 
cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
gdd400_all_new<-left_join(gdd400_all, cpp_year, by=c("pop"))
gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)

@@

We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions. Insect phenology studies often estimate generation times using growing degree day (GDD) models, which take advantage of the fact that insect development time is linearly temperature-dependent under moderate climate conditions. Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 growing degree-days. We used climate records over the past 20 years from our sampling locations to calculate a yearly GDD-based "deadline" for diapause at each location: the date after which eggs laid would not have enough growing degree days left in the year to complete another generation.

@@

```{r echo=FALSE}
kable(rbind(prettify(anova(model2)), prettify(anova(model1)), prettify(anova(model4)), prettify(anova(model3))), digits=3)  
```

```{r echo=FALSE}
prev_year_plots
```

```{r echo=FALSE}
kable(prettify(anova(all_prevyear_model)), digits=3) 
kable(prettify(anova(all_previnterval_model)), digits=3)

@@

The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process. Based on the distribution of percent differences from the initial counts (Figure S1A), we considered oviposition papers with more than 100% or less than 60% of the initial count to be less reliable data points; these papers made up less than 10% of both individual papers and total eggs quantified in the experiment. Excluding these data points resulted in 5 cages out of 286 in the total data set (1.7%) having less than 100 remaining eggs, so they were excluded from the photoperiod response curves. These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations (Figure S2). The final photoperiod response curves are shown in Figure S3.

@@

```{r, echo=FALSE}
DI_plot 
```
```{r echo=FALSE, warning=FALSE}
field_data_plot
```

```{r echo=FALSE}
gdd_deadline_plots
```
```{r echo=FALSE}
plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"), labels="AUTO")

```{r echo=FALSE, warning=FALSE}
gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd400_mean, fill=country, shape=interval)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Period") + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) +
  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Period") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=12)) + xlab("Latitude") + ylab("Mean GDD Deadline")

### 2020-02-20 bioclim_data.Rmd
Bioclim

```{r}
#Download worldclim data 
clim = getData('worldclim', var='bio', res=2.5) 
gain(clim) = 0.1
```

```{r}
#look up SpatialPoints 
# https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim
```

>>>>

```{r}
#Download worldclim data 
clim = getData('worldclim', var='bio', res=2.5) 
gain(clim) = 0.1 #values in bioclim dataset are scaled by 10
```

```{r}
# https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim

coords <- data.frame(x=c(-80.2, -80.4, -80.9, -81.8, -81.5, -80, -78.1, -78.4, -77.1, -77.4, -74.9, -74.1, 127.8, 130.6, 130.6, 130.3, 131.5, 135.8, 136, 136.7, 139, 140.9),
                     y=c(26.67, 27.58, 28.85, 30.33, 31.13, 33.75, 35.03, 36.35, 37.05, 38.63, 39.77, 40.73, 26.2, 31.6, 32.65, 33.25, 34.18, 35, 35.35, 36.56, 37.91, 38.29))
points <- SpatialPoints(coords)
values <- extract(clim,points)

```
The bioclimate variables used were: minimum temperature of coldest month
